\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-11-06 demystify.tex
Updates from Overleaf

There are various guides on writing statement, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available. So I will not talk too much about statements. In short, talk about your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it. Also, use the statement to talk about stuff that admission committee members might not know about, e.g., your Github project with 1K+ stars or  your regular contributions to well-known open-source projects. Finally, talk about what makes you different than the other 1000 applicants that we are reviewing (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).

@@

TA is common in the beginning when you haven't found your adviser who would pay you RA. As a TA, you spend up to 20 hrs/week and help professors with their classes (e.g., grading or teaching labs/recitation). Your TA is paid through the school/department, i.e., they hire you to help teach.  During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding typically is not available during the summer, which has much fewer courses.

\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically a default thing for good CS PhD programs. When you apply to be a full-time student,  state that you need financial assistant. It is common that the PhD committee will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.

\begin{commentbox}[Vu:]
At GMU CS, students admitted with TA have  4 years of GTA guaranteed and also receive  stipend for the \textbf{first} summer.
\end{commentbox}

@@

\subsection{Will I be miserable during my a PhD?}\label{sec:happy}
Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable.  Indeed there are issues with PhD and academic in general, but then there are also issues with jobs, family, and life.  Certainly there are bad relationships with professors, co-authors, and officemates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student), and it is good to be aware of those.  However, in reality there are many good mentors, supportive labmates and department, happy students, etc.  So don't let social media make you feel pessimistic and destroy your quest to advance knowledge.

@@

This section lists various cultural issues that international students should pay attention to.

@@

>>>>

There are various guides on writing statement, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available. So I will not talk too much about statements. In short, talk about your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it. Also, use the statement to talk about stuff that admission committee members might not know about, e.g., your Github project with 1K+ stars or  your regular contributions to well-known open-source projects. Finally, talk about what makes you different than the other applicants (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).

@@

TA is common in the beginning when you haven't found an adviser who would pay you RA. As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation). Your TA is paid through the school/department, i.e., they hire you to help teach.  During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding is not typically available during the summer, which has few courses.

\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically default for good CS PhD programs. When you apply to be a full-time student,  state that you need financial assistant. It is common that the PhD committee will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.

\begin{commentbox}[Vu:]
At GMU CS, students admitted with TA have  4 years of GTA guaranteed, and in some cases also receive  stipend for the first summer.
\end{commentbox}

@@

\subsection{Will I be miserable during my a PhD?}\label{sec:happy}
Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable. Certainly there are bad relationships with professors, co-authors, and lab mates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student), and it is good to be aware of those.  However, in reality there are many good mentors, supportive lab mates and department, happy students, etc.  So don't let social media make you feel pessimistic and deter your quest to advance knowledge.

@@

This section lists various cultural issues that international students might want to pay attention to.

@@

\paragraph{Gifts} In many countries, it is customary to give professors gifts, often during holidays.  These gifts can be costly and profs. sometimes expected them. In the US, it is uncommon and perfectly OK to not to give gifts. \emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated. Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences). In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.

There's a misconception that in the US it is all business, with professors as bosses who pay students for their work, and doing nice thing implies you expect something in return. However, the reality is quite the opposite. While people can be straightforward and direct, they are also friendly and informal. You can call your professor by their first name (\S\ref{sec:address}), disagree and argue with that (and can actually gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.  Many students and professors also maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
